---
title: "gtsummary in quarto"
format: html
---

```{r}
#| include: false

#output false and message false functionally the same here
#can also use include false to hide the whole chunk, both the input and output

library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
#| label: tbl-one
#| tbl-cap: Summary Statistics
#| echo: false

#need to store the table as an object in order to reference it in inline text
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))

#need to print the table again, since storing it as an object means it doesn't automatically print
table1

hisp_m_num <- inline_text(table1, variable = "race_eth_cat", level="Hispanic", column = "Male", pattern = "{n}")

hisp_f_num <- inline_text(table1, variable = "race_eth_cat", level="Hispanic", column = "Female", pattern = "{n}")

avg_wkdy_sleep_Hisp1 <- 
	filter(nlsy, race_eth_cat == "Hispanic") %>% 
	summarize (hispmean = mean(sleep_wkdy, na.rm = TRUE)) 


round(avg_wkdy_sleep_Hisp1$hispmean,2)

#OR

avg_wkdy_sleep_Hisp2 <- 
	filter(nlsy, race_eth_cat == "Hispanic") %>% 
	summarize (hispmean = round(mean(sleep_wkdy, na.rm = TRUE),2))

avg_wkdy_sleep_Hisp2$hispmean

#scales package --> number --> gives a way to specify decimales. This is giving 6.6 because the second decimal is a zero

avg_wkdy_sleep_Hisp1$hispmean
#6.601846


avg_wkdy_sleep <- round(mean(nlsy$sleep_wkdy, na.rm = TRUE),2)

avg_wkdy_sleep


```

@tbl-one is a standard Table 1 of summary statistics. There are `r hisp_m_num` Hispanic males and `r hisp_f_num` Hispanic females. On average, Hispanic participants slept for `r round(avg_wkdy_sleep_Hisp1$hispmean,2)` hours on weekdays, as compared to the average hours of weekday sleep among all participants, `r avg_wkdy_sleep`.

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing") |> 
  add_p(test = list(all_continuous() ~ "t.test", 
                    all_categorical() ~ "chisq.test")) |> 
  add_overall(col_label = "**Total**") |> 
  bold_labels() |> 
  modify_footnote(update = everything() ~ NA) |> 
  modify_header(label = "**Variable**", p.value = "**P**")
```

# Univariate regression

```{r}
tbl_uvregression(
  nlsy, 
  y = income,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, income, age_bir),
  method = lm)
```

```{r}
tbl_uvregression(
  nlsy, 
  y = glasses,
  include = c(sex_cat, race_eth_cat,
              eyesight_cat, glasses, age_bir),
  method = glm,
  method.args = list(family = binomial()),
  exponentiate = TRUE)
```

## Multivariable regressions

## Some regressions

```{r}
linear_model <- lm(income ~ sex_cat + age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r}
linear_model_int <- lm(income ~ sex_cat*age_bir + race_eth_cat, 
                   data = nlsy)
```

```{r}
logistic_model <- glm(glasses ~ eyesight_cat + sex_cat + income, 
                      data = nlsy, family = binomial())
```

## Tables

```{r}
tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))
```

```{r}
tbl_regression(
  logistic_model, 
  exponentiate = TRUE,
  label = list(
    sex_cat ~ "Sex",
    eyesight_cat ~ "Eyesight",
    income ~ "Income"
  ))
```

```{r}
tbl_no_int <- tbl_regression(
  linear_model, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth"
  ))

tbl_int <- tbl_regression(
  linear_model_int, 
  intercept = TRUE,
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    age_bir ~ "Age at first birth",
    `sex_cat:age_bir` ~ "Sex/age interaction"
  ))
```

```{r}
tbl_merge(list(tbl_no_int, tbl_int), 
          tab_spanner = c("**Model 1**", "**Model 2**"))
```
